Equation-free analysis of a dynamically evolving multigraph

### 2016-02-11 manuscript-materials/tex/multigraph-manuscript.tex
Everything works, manuscript nearly finalized.
All necessary figure-plotting routines moved to
./manuscript-materials/paper-figs.py, including advice on how to
generate the proper dataset.
Submitted version of manuscript

\documentclass[epjST, final]{svjour}
\usepackage{graphicx, subcaption, setspace, indentfirst, amsfonts, amsthm, empheq}
% \usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
% \usepackage[center]{titlesec}
\graphicspath{ {./figs/paper/} }
% \pagestyle{plain}
% \renewcommand\thesection{\Roman{section}}
% \renewcommand\thesubsection{\alph{subsection}.}
\newcommand\ignore[1]{}
% \setlength{\parskip}{0ex}
% \def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
% \let\endchangemargin=\endlist

@@

\abstract{In order to adapt traditional techniques from numerical
  analysis to complex network systems, we apply the equation-free
  framework to a dynamically evolving multigraph system. This
  approach is based on the coupling of short intervals of simulation
  with appropriately-defined lifting and restriction operators that
  map the full network description to suitable macroscopic
  variables, enabling both the acceleration of direct simulations
  through coarse-projective integration, and the identification of
  stationary states via a Newton-GMRES method. Additionally, we
  employ the dimensionality-reduction algorithms principal component
  analysis (PCA) and diffusion maps (DMAPS) to uncover hidden
  structure in the model.}

@@

Over the past decades, myriad systems have been formulated as complex
evolving networks, from highway traffic
\cite{joubert_large-scale_2010} to brain connectivity
\cite{hermundstad_learning_2011}. When modeling such systems, dynamics
are typically defined at a very fine scale, specifying individual node
and edge interactions; explicit closed equations governing macroscopic
properties are often unavailable
\cite{durrett_graph_2012,joubert_large-scale_2010,roche_agent-based_2011,swaminathan_modeling_1998}. These
interactions are often complicated functions dependent on multiple
system parameters, which, when coupled with the large network sizes
inherent in many interesting problems, makes numerical investigation
computationally prohibitive. Additionally, the lack of such explicitly
macroscopic equations precludes the use of traditional numerical
techniques such as optimization and fixed-point finding that could
offer valuable insight into the network's behavior, leaving little
alternative but to work with the full network system. Faced with this
dilemma, investigators must either restrict their attention to a
smaller, heuristically-motivated parameter space
\cite{hodgkin_quantitative_1952} or simplify the model, potentially
removing important features \cite{brown_variability_1999}. \par

Equation-free modeling offers a solution to these obstacles, allowing
one to investigate the complex network at a macroscopic level while
retaining full system details \cite{kevrekidis_equation-free:_2004,gear_equation-free_2003}. Underlying this method is the
assumption that, although we cannot analytically derive equations
governing network evolution, such equations do exist. Furthermore,
these implicit equations are assumed to be functions of only a few
dominating variables. Thus, the state of the complete network is in
fact, by this assumption, well described by these select
quantities. This may seem an onerous restriction, yet it is exactly
the behavior witnessed across network types: despite the complexity of
the full system, certain network properties evolve smoothly in time,
while the evolution of other, secondary properties, is often
correlated to that of a few primary variables
\cite{bold_equation-free_2014,rajendran_coarse_2011,siettos_equation-free_2011}. Once
these significant variables are uncovered, we can combine short
intervals of full system simulation with operators that map the full
system description to and from their representative coarse variables
to perform previously-unfeasible system-level analysis (see Section
\ref{sec:ef} for further details). \par

Here, we apply this framework to a dynamically evolving multigraph model. This offers a test of the methodology in the context of multigraphs, a previously unexplored case. We demonstrate the acceleration of network simulations through coarse-projective integration and the location of stationary states through a matrix-free Newton-GMRES method. In addition, principal component analysis and diffusion maps, two established dimensionality-reduction techniques, are shown to provide an algorithmic method of determining the underlying dimensionality of the system. \par

We begin in Section \ref{sec:m} with a description of the
model. Section \ref{sec:ef} provides further details of the
equation-free method, specifies how it was applied to our system, and
presents subsequent results. Section \ref{sec:dr} summarizes the
operation of PCA and DMAPS, and assesses their effectiveness in
analyzing hidden, low-dimensional structure in this network system.

@@

We study the edge-conservative preferential attachment model, a
detailed mathematical analysis of which can be found in
\cite{rath_time_2012} and \cite{rath_multigraph_2012}. The system
evolves in discrete steps $t = 0,1,\ldots t_f$, and we denote the
$n$-vertex graph at each point by $G_n(t)$. The initial state,
$G_n(0)$, is an arbitrary distribution of $m$ edges among the $n$
vertices. No restrictions are placed on this initial distribution:
multiple edges and loops are permitted.\ignore{an Erd\H{o}s-R\'{e}nyi
  random graph on $n$ vertices and $m$ edges. Here, however, $m$ may
  be larger than the total number of edges in a completely-connected
  $n$-vertex graph so we cannot assign a probability to each edges
  existence, instead placing $m$ edges at random in the graph.} The
system is then advanced step-by-step based on the following procedure:

\begin{enumerate}
\item Choose an edge $e_{ij} \in E(G)$ uniformly at random and flip a coin to label one of the ends as $v_{i}$
\item Choose a vertex $v_{k}$ using linear preferential attachment: $P(v_{k} = v_{l}) = \frac{d_{l} + \kappa}{\sum\limits_{i=1}^{n} d_{i} + n \kappa}$
\item Replace $e_{ij}$ with $e_{ik}$
\end{enumerate}

\noindent where $d_i$ is the degree of vertex $v_i$, $E(G)$ is the set of edges in the graph, and $\kappa \in (0, \infty)$ is a model parameter affecting the influence degrees have on the probability distribution. That is, taking $\lim\limits{\kappa \rightarrow 0}$, we recover ``pure'' preferential attachment, and probability scales directly with degree, while $\lim\limits_{\kappa \rightarrow \infty} P(v_k = v_l) = \frac{1}{n} \; \forall \; l$, and individual degrees have no effect. \par

Evolving the system in this manner, the degree sequence approaches a
stationary distribution over $O(n^3)$ steps. As explained in
\cite{rath_time_2012}, this distribution is dependent only on system
parameters $\rho = \frac{2*m}{n}$ and $\kappa$. Fig. (\ref{fig:dse})
illustrates the evolution of the degree sequence of two networks with
different initial configurations but equivalent values of $\rho$ and
$\kappa$; as expected, we observe they approach an identical stationary state.

@@

\begin{figure}[th!]
  \vspace{-5mm}
  \centering
  \begin{subfigure}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{lopsided-degs-a}
    \subcaption{\label{fig:lopsided-init}}
  \end{subfigure} %
  \begin{subfigure}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{erdos-degs-a}
    \subcaption{\label{fig:erdos-init}}
  \end{subfigure}
  \caption{Multigraph evolution dynamics for an $n=100$ vertex graph
    observed through the degree distribution. Two distinct transients
    are shown, initialized with (a) an Erd\H{o}s-R\'{e}nyi random
    graph with $m = 5,050$ edges and (b) a graph in which half the
    vertices are isolated, and half completely connected among
    themselves. \label{fig:dse}}
\end{figure}

\begin{figure}[th!]
  % \vspace{-5mm}
  % \centering
  % \begin{subfigure}[t]{0.49\textwidth}
  %   \centering
  %   \includegraphics[width=\textwidth]{triangle-slaving}
  %   \subcaption{\label{fig:sv-n2}}
  % \end{subfigure} %
  % \begin{subfigure}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{triangle-slaving-n2-n3}
    % \subcaption{\label{fig:sv-n3}}
  % \end{subfigure}
  \caption{Evolution of higher-order statistics (here, the total
    triangle count) for three different network initializations. They
    are quickly ($O(n^2)$ steps) drawn to a slow manifold on which
    they slowly evolve ($O(n^3)$ steps) to a steady
    state. \label{fig:sv}}
\end{figure}

@@

The first task in the implementation of the equation-free framework is to uncover those few, select variables that suffice to describing the behavior of the full, detailed network. This set of variables should be of lower dimensionality than the full system, and should evolve smoothly. Based on the profiles depicted in Fig. (\ref{fig:dse}), the graph's degree sequence, or equivalently degree distribution, is a good candidate. It progresses smoothly, and captures much information about the underlying network, while providing significant savings in dimensionality from an $O(n^2)$ adjacency matrix to a length-$n$ vector of degrees. \par

However, it is crucial to test whether the evolution of other
properties of the graph can be correlated to degree sequence. If not,
our current description is incomplete and there exist other important
coarse variables that must be accounted for in our macroscopic system
description. To assess this, we constructed graphs with identical
degree sequences but varying triangle counts and observed the
evolution of this property under the dynamics prescribed in Section
\ref{sec:m}. Fig. (\ref{fig:sv}) shows that, within a short time, the
triangle counts are drawn to a shared slow manifold, despite the varied
initial conditions. This suggests that the degree sequence is indeed
an adequate coarse variable for the system. \par

@@

\begin{figure}[ht!]
  \centering
  \includegraphics[width=\textwidth]{cpi-diagram}
  \caption{Schematic of our restriction ($R$) and lifting ($L$)
    procedures. The graph's sorted degree sequence is calculated
    ($R_1$) and then fit to a third-order polynomial ($R_2$). This
    maps graphs ($G$) to vectors of four coefficients
    ($\mathbf{c}$). To lift, the polynomial is used to generate a
    degree sequence ($L_1$), which is then used to construct a
    consistent graph ($L_2$). This process maps vectors of
    four coefficients ($\mathbf{c}$) to graphs
    ($G$). $\mathbf{R} = R_2 \circ R_1 $ is our restriction and $\mathbf{L} = L_2 \circ
    L_1$ is our lifting. \label{fig:cpi-diagram}}
\end{figure}

@@

The restriction operator translates the full network into its
equivalent coarse variables. Here, this involves mapping a multigraph
to its sorted degree sequence, a simple calculation. However, we can
continue to reduce the dimensionality of our coarse system by keeping
not the length-$n$ degree sequence, but a polynomial fitting of it. To
do so, we first sort the sequence to achieve a smooth, monotonically
increasing dataset, then fit the result to some function. Here, a
third-order polynomial was observed to closely approximate the
sequence throughout time; thus, our coarse description consisted of
the four polynomial coefficients specifying a particular sorted degree
sequence, as shown in Fig. (\ref{fig:cpi-diagram}). The restriction
operator therefore maps multigraphs to length-four coefficient
vectors. This procedure is represented by the blue arrows of
Fig. (\ref{fig:cpi-diagram}).

@@

The lifting operator performs the inverse role of restriction: translating coarse variables into full networks. Specifically in the context of our model, we map vectors of polynomial coefficients to full networks in a two-stage process. First, the coefficient vector is used to recreate a sorted degree sequence by evaluating the full polynomial at integer values and rounding to the nearest degree, as depicted in Fig. (\ref{fig:cpi-diagram}). If the sum of the resulting degree sequence is odd, a single degree is added to the largest value to ensure the sequence is graphical. Second, this degree sequence is used as input to a Havel-Hakimi algorithm, creating a graph whose degree sequence matches that of the input \cite{havel_remark_1955,hakimi_realizability_1962}. While the canonical Havel-Hakimi method produces simple graphs, it is not difficult to extend it to multigraphs by allowing the algorithm to wrap around the sequence, producing multiple edges and self-loops. The lifting procedure is represented by the red arrows of Fig. (\ref{fig:cpi-diagram}).

@@

The pieces described above are combined to accelerate simulations through coarse-projective integration. Denoting the lifting operator by $\mathbf{L} \; : \; c \rightarrow G$ where $c \in \mathbb{R}^4$ is the vector of polynomial coefficients, and the restriction operator as $\mathbf{R} \; : G \rightarrow c$, the method progresses as follows:

\begin{enumerate}
\item Advance the full system for $t_h$ steps
\label{cpi:heal}
\item Continue for $t_c$ steps, restricting the network at even intervals and saving the resulting coarse variables
\item Using the coarse variables collected over the previous $t_c$ steps, project each variable forward $t_p$ steps with a forward-Euler method
\label{cpi:proj}
\item With the new, projected coarse variables, lift to a full network
\label{cpi:init}
\item Repeat from Step (\ref{cpi:heal}) until desired number of steps have been reached
\end{enumerate}

Note that Step (\ref{cpi:heal}) is necessary to allow the system to approach its slow manifold. Upon initializing a new network in Step (\ref{cpi:init}), certain variables will be far from their expected values in a normal system; however, by hypothesis these quantities quickly evolve to a state in which they are strongly correlated with, or are functions of, the selected coarse variables. Thus, after a short interval of ``healing'', they are drawn to the expected trajectory, after which we begin sampling. This is analogous to Fig. (\ref{fig:sv}). The computational gains arise from the projective step, (\ref{cpi:proj}). Here, we advance the system $t_p$ steps at the cost of one evaluation of forward Euler, instead of $t_p$ direct evaluations. \\

Results of the application of this general framework with the specific lifting and restriction operators previously outlined are shown in Fig. (\ref{fig:cpi-results}). We see good agreement between the CPI-accelerated and normal systems, while reducing the number of detailed steps by one third. It is important to mention that this method was applied not to a single system instantiation, but to a large ensemble. This ensured that when properties such as the fitted polynomial coefficients were averaged over the ensemble they evolved smoothly despite the stochastic nature of the system.

\begin{figure}[h!]
  \vspace{-5mm}
  \centering
  \begin{subfigure}{0.59\textwidth}
    \centering
    \includegraphics[width=\textwidth]{cpi-3d-comp-stretch}
    \subcaption{\label{fig:cpi-error}}
  \end{subfigure} %
  \begin{subfigure}{0.39\textwidth}
    \centering
    \includegraphics[width=\textwidth]{cpi-relative-error-n3}
    \subcaption{\label{fig:self-error}}
  \end{subfigure}%
  \caption{CPI results: (a) the evolution of the CPI-accelerated
    degree sequence (red) compared to direct simulation (blue) and (b)
    error in cpi-accelerated runs calculated by comparing
    cpi-accelerated degree sequences to those arising from an ensemble
    of direct simulations. \label{fig:cpi-results}}
\end{figure}

\section{Newton-GMRES}

Aside from the computational savings of model simulations offered by
CPI, the equation free framework also permits the calculation of
system steady states through matrix-free Newton-GMRES
algorithms. Referring back to the CPI procedure outlined in
Sec. (\ref{sec:cpi}), we can define an operator
$\Theta: d \rightarrow d$ projecting coarse variables at $t$ to their
values at $t + \delta t$: $d(t+\delta t) =\Theta(d(t))$. Note that in
this section, we take the sorted degree sequence as our coarse
variable. A system fixed point could then be located by employing
Newton's method to solve the equation

\begin{align}
\label{eq:f}
F(d) = \Theta(d) - d = 0
\end{align}

However, this requires the repeated solution of
$DF(d^{(k)}) \delta d^{(k+1)} = -F(d^{(j)})$, in which the system
Jacobian, $DF$ is unavailable to us. Thankfully, we may circumvent
this obstacle by estimating the directional derivatives
$DF(d) \d dot \delta d$ via a difference approximation of the form

\begin{align}
  DF(d) \ddot \delta d \approx \frac{\| \delta d \| F(d + h \| d \| \frac{\delta d}{\| \delta d \|}) - F(d)}{h \| d \|}
\end{align}

for nonzero $d$, which in turn is evaluated through calls to $F$ as defined in Eq. (\ref{eq:f})

This is precisely the approach of the Newton-GMRES method, in which the solution to a linear system is searched for in expanding Krylov subspaces \cite{kelley_solving_2003}. Applying this algorithm in conjunction with the $\Theta$ operator defined in Sec. (\ref{sec:cpi}) allowed us to locate the stationary distribution without simply running and observing the full system for long times, as would otherwise be required. Results are shown in Fig. (\ref{fig:newton-results}).

\begin{figure}[h!]
  \vspace{-5mm}
  \centering
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{newton-error}
    \subcaption{\label{fig:newton-error}}
  \end{subfigure} %
  \begin{subfigure}{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{newton-comp}
    \subcaption{\label{fig:newton-comp}}
  \end{subfigure}%
  \caption{Coarse Newton-GMRES results: (a) evolution of error in the
    coarse Newton-GMRES scheme and (b) visual comparison of the
    algorithm's solution to the stationary state arising from direct
    simulation. \label{fig:newton-results}}
\end{figure}

>>>>

\documentclass[epjST, final]{svjour} \usepackage{graphicx, subcaption,
  setspace, indentfirst, amsfonts, amsthm, empheq}
% \usepackage[top=1in, bottom=1in, left=1in, right=1in]{geometry}
% \usepackage[center]{titlesec}
\graphicspath{ {../figs/} }
% \pagestyle{plain} \renewcommand\thesection{\Roman{section}}
% \renewcommand\thesubsection{\alph{subsection}.}
\newcommand\ignore[1]{}
% \setlength{\parskip}{0ex}
% \def\changemargin#1#2{\list{}{\rightmargin#2\leftmargin#1}\item[]}
% \let\endchangemargin=\endlist

@@

\abstract{In order to illustrate the adaptation of traditional
  continuum numerical techniques to the study of complex network
  systems, we use the equation-free framework to analyze a dynamically
  evolving multigraph.
%
  This approach is based on coupling short intervals of direct dynamic
  network simulation with appropriately-defined lifting and
  restriction operators, mapping the detailed network description to
  suitable macroscopic (coarse-grained) variables and back.
%
  This enables the acceleration of direct simulations through Coarse
  Projective Integration (CPI), as well as the identification of
  coarse stationary states via a Newton-GMRES method.
%
  We also demonstrate the use of data-mining, both linear (principal
  component analysis, PCA) and nonlinear (diffusion maps, DMAPS) to
  determine good macroscopic variables (observables) through which one
  can coarse-grain the model.}

@@

  Over the past decades, complex networks have been used as the model
  of choice for a truly broad class of physical problems, ranging from
  highway traffic \cite{joubert_large-scale_2010} to brain
  connectivity \cite{hermundstad_learning_2011}.
%
  When modeling such systems, dynamics are typically defined at a very
  detailed, ``fine'' scale, specifying individual node and edge
  interactions; explicit closed equations governing network
  macroscopic (collective) properties are often unavailable
  \cite{durrett_graph_2012,joubert_large-scale_2010,roche_agent-based_2011,swaminathan_modeling_1998}.
%
  The detailed interactions are often complicated functions dependent
  on multiple system parameters, which, when one also accounts for
  large network sizes inherent in many interesting problems, can make
  systematic numerical exploration computationally prohibitive.
%
  The lack of \textit{explicit} macroscopic equations prohibits the
  use of traditional numerical techniques such as fixed-point
  computation and stability analysis, that could offer valuable
  insight into the network's behavior, leaving little alternative but
  to work with full direct simulations of the entire network.
%
  Faced with these inherent limitations, investigators must either
  restrict their attention to a more modest parameter space
  \cite{hodgkin_quantitative_1952} or simplify the network model,
  potentially removing important features
  \cite{brown_variability_1999}. \par

  Equation-free modeling offers a promise to circumvent these
  challenges, allowing one to investigate the complex network at a
  macroscopic level while retaining the effects of full system details
  \cite{kevrekidis_equation-free:_2004,gear_equation-free_2003}.
%
  Underlying this method is the assumption that, although we cannot
  analytically derive equations governing network evolution, such
  closed equations do, in principle, exist.
%
  Furthermore, these unavailable equations are assumed to involve a
  small number of dominant collective (coarse-grained) variables.
%
  The important features of the complete network can be in fact, by
  this assumption, well represented by these select observable
  quantities.
%
  This may seem too restrictive, yet it is exactly the behavior
  witnessed across many network types: despite the initial complexity
  of the full system configuration, certain collective network
  properties appear to evolve smoothly in time, while the evolution of
  other, ``secondary'' properties, can be strongly correlated with
  that of the few ``primary'' variables
  \cite{bold_equation-free_2014,rajendran_coarse_2011,siettos_equation-free_2011}.
%
  Once these significant variables are uncovered, we can combine short
  intervals of full system simulation with operators that map the full
  system description to and from its representative coarse variable
  ``summary'', thus enabling previously-infeasible system-level
  analysis (see Section \ref{sec:ef} for further details). \par

  Here, we apply this framework to a dynamically evolving multigraph
  model.
%
  This offers a test of the methodology in the previously unexplored
  context of multigraphs.
%
  We demonstrate the acceleration of direct network simulations
  through Coarse Projective Integration (CPI), and the location of
  coarse stationary states through a matrix-free Newton-GMRES method.
%
  In addition, principal component analysis (PCA) and diffusion maps
  (DMAPS), two well established dimensionality-reduction techniques,
  are shown to enable an algorithm for characterization of the
  underlying low-dimensional behavior of the system. \par

  The paper is organized as follows: we begin in Section \ref{sec:m}
  with a description of the multigraph dynamic model.
%
  Sections \ref{sec:ef} and \ref{sec:ng} provide most details of the
  equation-free approach, specify how it was applied to our system,
  and present subsequent results.
%
  Section \ref{sec:dr} summarizes our use of PCA and DMAPS, and
  assesses their effectiveness in analyzing hidden, low-dimensional
  structure in this network system.

@@

  We study the edge-conservative preferential attachment model, a
  detailed mathematical analysis of which can be found in
  \cite{rath_time_2012} and \cite{rath_multigraph_2012}.
%
  The system evolves in discrete steps $t = 0,1,\ldots t_f$, and we
  denote the $n$-vertex graph at each point by $G_n(t)$.
%
  The initial state, $G_n(0)$, is an arbitrary distribution of $m$
  edges among the $n$ vertices; the total number of both edges and
  vertices is conserved.
%
  No restrictions are placed on this initial distribution: multiple
  edges and loops are permitted.
  % \ignore{an Erd\H{o}s-R\'{e}nyi random graph on $n$ vertices and
  % $m$ edges. Here, however, $m$ may be larger than the total number
  % of edges in a completely-connected $n$-vertex graph so we cannot
  % assign a probability to each edges existence, instead placing $m$
  % edges at random in the graph.}
  The system is then advanced step-by-step based on the following
  procedure:

  \begin{enumerate}
  \item Choose an edge $e_{ij} \in E(G)$ uniformly at random, and flip
    a coin to label one of the ends as $v_{i}$
  \item Choose a vertex $v_{k}$ using linear preferential attachment:
    $P(v_{k} = v_{l}) = \frac{d_{l} + \kappa}{\sum\limits_{i=1}^{n}
      d_{i} + n \kappa}$
    \label{step:prob}
  \item Replace $e_{ij}$ with $e_{ik}$,
  \end{enumerate}

  \noindent where $d_i$ is the degree of vertex $v_i$, $E(G)$ is the
  set of edges in the graph, and $\kappa \in (0, \infty)$ is a model
  parameter affecting the influence degrees have on the probability
  distribution in Step \ref{step:prob}.
%
  That is, taking $\lim\limits{\kappa \rightarrow 0}$, we recover
  ``pure'' preferential attachment, and probability scales directly
  with degree,
  %%% YGK immediately? after a short time ? at steady state ?  if at
  %%% steady state say it ?  ATH This refers to the dynamics at all
  %%% times. At every step, we need some probability of choosing a
  %%% given vertex v_i: if k -> 0, this probability is proportional to
  %%% deg(v_i), and if k -> infinity this probability is uniform over
  %%% all vertices (and degree has no effect. Valid at every step.
  while
  $\lim\limits_{\kappa \rightarrow \infty} P(v_k = v_l) = \frac{1}{n}
  \; \forall \; l$, and individual degrees have no effect.
  %%% YGK you mean individual initial degrees ?
%%%
  \par

  Evolving the system in this manner, the degree sequence approaches a
  stationary distribution over $O(n^3)$ steps.
%
  As explained in \cite{rath_time_2012}, this distribution is
  dependent only on the system parameters $\rho = \frac{2*m}{n}$ and
  $\kappa$.
%
  Fig. (\ref{fig:dse}) illustrates the evolution of the degree
  sequence of two networks with different initial configurations but
  the same values of $\rho$ and $\kappa$ respectively; as expected, we
  observe they approach an identical stationary state.

@@

  \begin{figure}[th!]
    \vspace{-5mm} \centering
    \begin{subfigure}[t]{0.49\textwidth}
      \centering
      \includegraphics[width=\textwidth]{lopsided-degs-a}
      \subcaption{\label{fig:lopsided-init}}
    \end{subfigure} %
    \begin{subfigure}[t]{0.49\textwidth}
      \centering
      \includegraphics[width=\textwidth]{erdos-degs-a}
      \subcaption{\label{fig:erdos-init}}
    \end{subfigure}
    \caption{Multigraph evolution dynamics for an $n=100$ vertex graph
      observed through the degree sequence (after sorting the nodes by
      degree). Two distinct transients are shown, initialized with (a)
      an Erd\H{o}s-R\'{e}nyi random graph with $m = 5,050$ edges and (b)
      a graph in which half the vertices are isolated, and half
      uniformly share $m = 5,050$ edges. Both approach the same ultimate
      stationary sequence. \label{fig:dse}}
  \end{figure}

  \begin{figure}[th!]
    % \vspace{-5mm}
    % \centering
    % \begin{subfigure}[t]{0.49\textwidth}
    %   \centering
    %   \includegraphics[width=\textwidth]{triangle-slaving}
    %   \subcaption{\label{fig:sv-n2}}
    % \end{subfigure} %
    % \begin{subfigure}[t]{0.49\textwidth}
    \centering
    \includegraphics[width=0.8\textwidth]{triangle-slaving-n2-n3}
    % \subcaption{\label{fig:sv-n3}}
    % \end{subfigure}
    \caption{Evolution of higher-order network statistics (here, the
      total triangle count) for three different network
      initializations. They are quickly ($O(n^2)$ steps) drawn to a slow
      manifold on which they slowly evolve over ($O(n^3)$ steps) to a
      stationary state. \label{fig:sv}}
  \end{figure}
  %%% 
  %%% YGK maybe you can draw a short horizontal broken line at the "end"
  %%% of the graph, where "presumably" the steady state value is
  %%% extending from the boundary of the plot "inward" in the plot maybe
  %%% a thick black broken line ?

@@

  A prerequisite to the implementation of equation-free algorithms is
  the determination of those few, select variables with which one can
  ``close'' a useful approximate description of the coarse-grained
  dynamics of the full, detailed network.
  % 
  This set of variables should clearly be (much) less than those of the
  full system state, and they should evolve smoothly (so, either our
  network would have a large number of nodes, or we would take
  expectations over several realizations of networks with the same
  collective features).
  % 
  Based on the profiles depicted in Fig. (\ref{fig:dse}), the graph's
  degree sequence, (equally informatively, its degree distribution) is a
  good candidate collective observable.
  % 
  It appears to evolve smoothly, while also providing significant
  savings in dimensionality, from an $O(n^2)$ adjacency matrix to a
  length-$n$ vector of degrees. \par

  It now becomes crucial to test whether the evolution of other
  properties of the graph can be correlated to the degree sequence.
  % 
  If not, our current description is incomplete (an equation cannot
  close with only the degree sequence) and there exist other important
  coarse variables that must be included in our macroscopic system
  description.
  % 
  To assess this, we constructed graphs with identical degree sequences
  but varying triangle counts and recorded the evolution of this
  observable under the dynamics prescribed in Section
  \ref{sec:m}. Fig. (\ref{fig:sv}) shows that, within a short time, the
  triangle counts are drawn to an apparent shared ``slow manifold'',
  despite the variation in initial conditions.
  % 
  This supports our selection of the degree sequence as an adequate
  coarse variable to model the system. \par

@@

  \begin{figure}[ht!]
    \centering
    \includegraphics[width=\textwidth]{cpi-diagram}
    \caption{Schematic of our restriction ($\mathbf{R}$) and lifting
      ($\mathbf{L}$) procedures. The graph's sorted degree sequence is
      calculated ($R_1$) and then fit to a third-order polynomial
      ($R_2$). This maps graphs ($G$) to vectors of four coefficients
      ($\mathbf{c}$). To lift, the polynomial is used to generate a
      degree sequence ($L_1$), which is then used to construct a
      consistent graph ($L_2$). This process maps vectors of four
      coefficients ($\mathbf{c}$) to graphs ($G$).
      $\mathbf{R} = R_2 \circ R_1 $ is our restriction and
      $\mathbf{L} = L_2 \circ L_1$ is our
      lifting. \label{fig:cpi-diagram}}
  \end{figure}

@@

  The restriction operator ``translates'' the full network description
  to its corresponding coarse variable values.
  % 
  Here, this involves mapping a multigraph to its sorted degree
  sequence, a simple calculation.
  % 
  However, we may be able to further reduce the dimensionality of our
  coarse system by keeping not the entire length-$n$ degree sequence,
  but rather a low-order polynomial fitting of it.
  % 
  To do so, we first sort the sequence to achieve a smooth,
  monotonically increasing dataset, then fit the result to a third-order
  polynomial, which had been observed to closely approximate the
  sequence evolution throughout time.
  % 
  Our coarse description thus consisted, at every moment in time, of the
  four polynomial coefficient values, specifying a particular sorted
  degree sequence, as shown in Fig. (\ref{fig:cpi-diagram}).
  % 
  The restriction operator therefore maps multigraphs to length-four
  coefficient vectors.
  % 
  This procedure is represented by the blue arrows of
  Fig. (\ref{fig:cpi-diagram}).

@@

  The lifting operator performs the inverse role of restriction:
  translating coarse variables into full networks.
  % 
  This is clearly a one-to-many mapping.
  % 
  Specifically in the context of our model, we map vectors of polynomial
  coefficients to full networks in a two-stage process.
  % 
  First, the coefficient vector is used to recreate a sorted degree
  sequence by evaluating the full polynomial at integer values and
  rounding to the nearest degree, as depicted in
  Fig. (\ref{fig:cpi-diagram}).
  % 
  If the sum of the resulting degree sequence is odd, a single degree is
  added to the largest value to ensure the sequence is graphical.
  % 
  Second, this degree sequence is used as input to any algorithm that
  produces a network consistent with a graphical degree sequence (here
  we typically used a Havel-Hakimi algorithm, creating a graph whose
  degree sequence matches that of the input
  \cite{havel_remark_1955,hakimi_realizability_1962}).
  % 
  While the canonical Havel-Hakimi method produces simple graphs, it is
  not difficult to extend it to multigraphs by allowing the algorithm to
  wrap around the sequence, producing multiple edges and self-loops.
  %%% YGK is this you, Alexander, is it Balazs, did you read it
  %%% somewhere, is there a reference ?  ATH This is me, but Balazs
  %%% confirmed this in his comments when he said that Havel Hakimi was
  %%% even overly-complex for this case as there are even simpler
  %%% methods one could use. I think it is easy to prove.
  The lifting procedure is represented by the red arrows of
  Fig. (\ref{fig:cpi-diagram}).

@@

  The components described above are combined to accelerate simulations
  through Coarse Projective Integration.
  % 
  Denoting the lifting operator by $\mathbf{L} \; : \; c \rightarrow G$
  where $c \in \mathbb{R}^4$ is the vector of polynomial coefficients,
  and the restriction operator as $\mathbf{R} \; : G \rightarrow c$, the
  algorithm progresses as follows:

  \begin{enumerate}
  \item Advance the full system for $t_h$ steps
    \label{cpi:heal}
  \item Continue for $t_c$ steps, restricting the network at intervals
    and recording the coarse variable values
  \item Using the coarse variables collected over the previous $t_c$
    steps, project each variable forward $t_p$ steps with, here, a
    forward-Euler method
    \label{cpi:proj}
  \item With the new, projected coarse variables, lift to one (or more
    copies of) full network
    \label{cpi:init}
  \item Repeat from Step (\ref{cpi:heal}) until the desired time has
    been reached.
  \end{enumerate}

  Note that Step (\ref{cpi:heal}) is intended to allow a singularly
  perturbed quantity to approach its slow manifold.
  % 
  Upon initializing a new full system (here, a new network) in Step
  (\ref{cpi:init}), higher order observables (for example, the triangle
  count) will be far from the values they would attain in a detailed
  simulation.
  % 
  If the coarse system model closes with our chosen variables, then
  either (a) these quantities do not affect the dynamics of the degree
  sequence; or (b) they do, but then by hypothesis these quantities
  quickly evolve to functions of the selected coarse variables (i.e.,
  they approach the slow manifold).
  % 
  In the second case, after a short interval of ``healing'' they will be
  drawn to the expected trajectory, after which we begin sampling.
  % 
  This is analogous to Fig. (\ref{fig:sv}).
  % 
  The computational gains arise from the projective step,
  (\ref{cpi:proj}).
  % 
  Here, we advance the system $t_p$ steps at the cost of one evaluation of forward Euler, instead of $t_p$ direct detailed steps. \\
  %%% YGK do you say what your step counts are in your actual
  %%% computation ?  ATH added specifics into the paragraph below

  Results of the application of this general framework with the specific
  lifting and restriction operators previously outlined are shown in
  Fig. (\ref{fig:cpi-results}). We used an $n=100$-vertex graph with
  $m=50000$ edges and parameter value $\kappa=1$. We ran the model for a
  total of $10 \cdot n^3$ steps, with $t_h = 10 \cdot n^2$, $t_c = n^3$
  and $t_p = 50 \cdot n^2$.
  % 
  We see good agreement between the CPI-accelerated and normal systems,
  while reducing the number of detailed steps by one third.
  % 
  It is important to mention that this method was applied not to a
  single system instantiation,
  % t
  but to an ensemble of fifty such realizations.
  %%% YGK say how many ?
  %%% 
  %%% ATH done (fifty)
  This ensured that when properties such as the fitted polynomial
  coefficients were averaged over the ensemble they evolved smoothly
  despite the stochastic nature of the system; in effect, we are
  computing the ``expected network evolution'' averaged over consistent
  realizations.

  \begin{figure}[h!]
    \vspace{-5mm} \centering
    \begin{subfigure}{0.59\textwidth}
      \centering
      \includegraphics[width=\textwidth]{cpi-3d-comp-stretch}
      \subcaption{\label{fig:cpi-error}}
    \end{subfigure} %
    \begin{subfigure}{0.39\textwidth}
      \centering
      \includegraphics[width=\textwidth]{cpi-relative-error-n3}
      \subcaption{\label{fig:self-error}}
    \end{subfigure}%
    \caption{CPI results: (a) the evolution of the CPI-accelerated
      degree sequence (red) compared to direct simulation (blue) and (b)
      error in CPI-accelerated runs calculated by comparing
      CPI-accelerated degree sequences to those arising from an ensemble
      of direct simulations. \label{fig:cpi-results}}
  \end{figure}

  \section{Coarse Newton-GMRES}
  \label{sec:ng}
  Aside from the computational savings of model simulations offered by
  CPI, the equation free framework also permits the calculation of
  system steady states through fixed point (here, matrix-free
  Newton-GMRES) algorithms.
  % 
  Referring back to the CPI procedure outlined in Sec. (\ref{sec:cpi}),
  we can define an operator $\Theta: \mathbf{d} \rightarrow \mathbf{d}$
  projecting coarse variables at $t$ to their values at $t + \delta t$:
  $\mathbf{d}(t+\delta t) =\Theta(\mathbf{d}(t))$.
  % 
  Note that in this section, we take the sorted degree sequence as our
  coarse variable.
  % 
  A system fixed point could then be located by employing Newton's
  method to solve the equation

  \begin{align}
    \label{eq:f}
    \mathbf{F}(\mathbf{d}) \equiv \Theta(\mathbf{d}) - \mathbf{d} = 0
  \end{align}

  %%% YGK can you consistently make F and d bold, so that they are
  %%% vectors ?
  %%% 
  %%% ATH done
